---
title: "Data Analysis Project"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```



# Student Information

**Name**: Manasi Dhananjay Choughule 





# Introduction

This semester we will be working with a dataset of auto claims filed by customers of an automobile insurance company located in the southwest and western regions of the United States.

Insurance companies depend on accurate pricing estimates to maintain profitability. Auto policies must be priced so that the insurance company makes a profit in the long run, given the costs of their customers' payouts for accident repairs, total loss car replacements, medical fees, and legal fees.

The executives at this insurance company have noticed declining profitability over the last several years and have hired you as a data science consultant to evaluate their claims data and make recommendations on pricing, customer behavior, and car insurance policy adjustments.

The objective of this project is to perform an exploratory data analysis on the claims_df dataset and produce an executive summary of your key insights and recommendations to the executive team at the insurance company.

Before you begin, take a moment to read through the following insurance company terms to familiarize yourself with the industry:  [Auto Insurance
Terms](https://www.iii.org/article/auto-insurance-jargon-buster "Auto Insurance Terms"){target="_blank"} 



## Auto Claims Data

The `claims_df` data frame is loaded below and consists of 6,249 auto claims submitted by customers of the insurance company. The rows in this data frame represent a single claim with all of the associated features that are displayed in the table below.

**Note**: If you have not installed the `tidyverse` package, please do so by going to the `Packages` tab in the lower right section of RStudio, select the `Install` button and type `tidyverse` into the prompt. If you cannot load the data, then try downloading the latest version of R (at least 4.0). The `readRDS()` function has different behavior in older versions of `R` and may cause loading issues.


```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

claims_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/claims_df.rds'))

```


### Raw Data

```{r}
claims_df
```
<br>
<br>


# Exploratory Data Analysis


Executives at this company have hired you as a data science consultant to evaluate their claims data and make recommendations on pricing, customer behavior, and car insurance policy adjustments.

You must think of **at least 8 relevant questions** that will provide evidence for your recommendations.

The goal of your analysis should be discovering which variables drive the differences between customers with large lifetime values and
customers who cost the company more than they provide in revenue through monthly premiums.

Some of the many questions you can explore include:

-   Are there types of customers, based on their policy or demographics,
    that are highly profitable?

-   Do certain policies have a lower number of claims, leading to large
    profits?

-   Are there "problem customers" which have a large number of claims?

You must answer each question and provide supporting data summaries with either a summary data frame (using `dplyr`/`tidyr`) or a plot (using `ggplot`) or both.

In total, you must have a minimum of 5 plots and 4 summary data frames for the exploratory data analysis section. Among the plots you produce, you must have at least 4 different types (ex. box plot, bar chart, histogram, heat map, etc...)

Each question must be answered with supporting evidence from your tables and plots.


Analyze the dataset:
```{r}
print("Unique in customer state")
unique(claims_df$customer_state)
```
```{r}
print("Unique in highest_education")
unique(claims_df$highest_education)
```
```{r}

print("Unique in employment_status")
unique(claims_df$employment_status)
```
```{r}
print("Unique in gender")
unique(claims_df$gender)
```
```{r}
print("range in income")
min(claims_df$income)
max(claims_df$income)
mean(claims_df$income)
median(claims_df$income)
```
```{r}
print("Unique in residence_type")
unique(claims_df$residence_type)
```
```{r}
print("Unique in marital_status")
unique(claims_df$marital_status)
```
```{r}
print("Unique in sales_channel")
unique(claims_df$sales_channel)
```
```{r}
print("Unique in sales_channel")
unique(claims_df$sales_channel)
```
```{r}
print("Unique in coverage")
unique(claims_df$coverage)
```
```{r}
print("Unique in policy")
unique(claims_df$policy)
```
```{r}
print("Unique in vehicle_class")
unique(claims_df$vehicle_class)
```
```{r}
print("Unique in vehicle_size")
unique(claims_df$vehicle_size)
```
```{r}
print("range in monthly_premium")
min(claims_df$monthly_premium)
max(claims_df$monthly_premium)
mean(claims_df$monthly_premium)
median(claims_df$monthly_premium)
```
```{r}
print("range in months_policy_active")
min(claims_df$months_policy_active)
max(claims_df$months_policy_active)
mean(claims_df$months_policy_active)
median(claims_df$months_policy_active)
```
```{r}
print("range in months_since_last_claim")
min(claims_df$months_since_last_claim)
max(claims_df$months_since_last_claim)
mean(claims_df$months_since_last_claim)
median(claims_df$months_since_last_claim)
```
```{r}
print("range in current_claim_amount")
min(claims_df$current_claim_amount)
max(claims_df$current_claim_amount)
mean(claims_df$current_claim_amount)
median(claims_df$current_claim_amount)
```
```{r}
print("range in customer_lifetime_value")
min(claims_df$customer_lifetime_value)
max(claims_df$customer_lifetime_value)
mean(claims_df$customer_lifetime_value)
median(claims_df$customer_lifetime_value)
```
## Question 1

**Question**:
Does gender and marital status affect the claim rate? 

**Answer**:
The gender does not affect the claim rate but married clients have significantly more claim rate.   


```{r}
ggplot(claims_df, aes(x=marital_status, y=total_claims, fill = gender))+
  geom_col()+
  ggtitle("Marital Status and Gender V/s Total claims")

```

## Question 2



**Question**:Does any state has more number of claims?

**Answer**:
California has significantly higher number of claims. It is followed by Oregon, Arizona, Nevada, and Washington.

```{r}
claims_with_state <- claims_df[c(2,18)]
claims_with_state_summarized <- claims_with_state %>%
                                  group_by(customer_state) %>%
                                  summarize(total_claims= sum(total_claims)) 
claims_with_state_summarized

ggplot(claims_with_state_summarized, aes(x="", y=total_claims, fill=customer_state)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  theme_void() 

```
## Question 3



**Question**: Are there types of customers, based on their policy or demographics, that are highly profitable?


**Answer**:
On calculating the total amount paid by the customer, we can find the percentage profit per customer. For this analysis I have only considered the customers that gave at least 50% profit to the company. Further, we will call these customers "Highly profitable customers." There are 1,148 highly profitable customers. In all the states employed customers bring most profit. Also, personal policy type has most number of highly profitable customers. Extended coverage is also profitable. 



```{r}

claims_df_high_profit <- claims_df %>%
                          filter(customer_lifetime_value>0) %>%
                          mutate(percentage_profit = ((monthly_premium * months_policy_active)-total_claims_amount)/(monthly_premium *     months_policy_active)*100) %>%
  arrange(percentage_profit) %>%
  filter(percentage_profit>50)

claims_df_high_profit 
claims_df_high_profit%>% group_by(customer_state) %>%count(employment_status)
claims_df_high_profit%>% group_by(policy) %>%count(policy) 
claims_df_high_profit%>% group_by(coverage) %>%count(coverage)
```
## Question 4



**Question**: Is there any specific trend between education, residential type, and income of the highly profitable customers?



**Answer**: Most of the highly profitable customers have completed high schools, associate degree, and bachelor degree. Profitability is significantly low for master degree and doctoral. For income between 10k to 70k is the range where we get most of the profit. Most of the highly profitable customers are from suburban.

```{r fig.align="center", echo = FALSE,fig.width = 14}
claims_df_high_profit_education <- claims_df_high_profit%>% group_by(customer_state) %>%count(highest_education) 
claims_df_high_profit_education
ggplot(claims_df_high_profit_education, aes(x=highest_education, y=n, fill = highest_education))+
  geom_col()+
  facet_grid(~customer_state)+
  ggtitle("Highly profitable customers' education in each state ") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r}
claims_df_high_profit <- claims_df_high_profit%>%mutate(gross_income = round(income/10000))

ggplot(claims_df_high_profit, aes(x=factor(gross_income))) +
  geom_histogram(stat="count", fill="blue")+
  ggtitle("Highly profitable customers' income distriution ")

```
```{r}
claims_df_high_profit%>% group_by(residence_type) %>%count(residence_type) 
```

## Question 5



**Question**: Are there "problem customers" which have a large number of claims?



**Answer**: Any customer with negative lifetime value (they received more money from the company than they paid to the insurance company), would be problem customer for the company. On filtering the customers with negative lifetime value and finding the median lifetime value for each number of claim made, we could see the total claim value paid to customers is increasing from 2 claims. 1005 customers have 3 total claims and 255 customers have 4 total claims. Moreover, there is drastic increase in median loss caused to the company as the total claims increases. 

```{r}
problem_customers <- claims_df %>%
                      filter(customer_lifetime_value<0) %>%
                      group_by(total_claims) %>%
                      mutate(problem_customer_avg_customer_lifetime_value = median(customer_lifetime_value)) %>%
                      count(total_claims, problem_customer_avg_customer_lifetime_value)
problem_customers

good_customers <- claims_df_high_profit%>%
                      group_by(total_claims) %>%
                      mutate(highly_profitable_avg_customer_lifetime_value = median(customer_lifetime_value)) %>%
                      count(total_claims, highly_profitable_avg_customer_lifetime_value)
good_customers

ggplot(problem_customers, aes(x= total_claims, y = problem_customer_avg_customer_lifetime_value))+ 
  geom_line()+
  ggtitle("Bad customers' claims with customer lifetime value")

ggplot(good_customers, aes(x= total_claims, y = highly_profitable_avg_customer_lifetime_value))+ 
  geom_line() + 
  ggtitle("Highly profitable customers' claims with customer lifetime value")
  
```


## Question 6



**Question**: Is there any car category that is causing more loss? Also any sales channel has more loss making customers?

**Answer**: Midsize car category is causing more loss for all sales channel. Four-door vehicle class is causing more loss. Also, Agent and branch are the sales channel that has more loss making customers.   



```{r fig.align="center", echo = FALSE,fig.width = 10}
loss_making_customers <- claims_df %>%
                      filter(customer_lifetime_value<0) 

vs <- loss_making_customers%>% group_by(vehicle_size) %>%count(sales_channel) 
vs

ggplot(vs, aes(x= sales_channel, y=n, color = sales_channel))+
  geom_point() +
  facet_grid(. ~ vehicle_size) +
  ggtitle("Sales channel and car category for loss making customers")
```
```{r}
loss_making_customers%>% group_by(vehicle_size) %>%count(vehicle_size) 
```
```{r}
loss_making_customers%>% group_by(vehicle_class) %>%count(vehicle_class) 
```

## Question 7

**Question**: Which vehicle class has maximum high profitable customers? Any sales channel bringing more high profitable customers?



**Answer**: Luxury suv and luxury cars are most profitable. All midsize car owners are profitable. Agents and branches are more profitable. 

```{r fig.align="center", echo = FALSE,fig.width = 10}
hp <- claims_df_high_profit%>% group_by(vehicle_size) %>%count(sales_channel) 
hp

ggplot(hp, aes(x= sales_channel, y=n, color = sales_channel))+
  geom_point() +
  facet_grid(. ~ vehicle_size) +
  ggtitle("Sales channel and car category for Highly Profitable customers")
```
```{r}
claims_df_high_profit%>% group_by(vehicle_size) %>%count(vehicle_size) 
```
```{r}
claims_df_high_profit%>% group_by(vehicle_class) %>%count(vehicle_class) 
```

```{r fig.align="center", echo = FALSE,fig.width = 10}
ggplot(claims_df_high_profit, aes(x= vehicle_class, y=percentage_profit, fill = vehicle_class))+
  geom_boxplot()+
  facet_grid(~vehicle_size)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("Vehicles v/s percentage profit")
```
## Question 8

**Question**: Are highly profitable clients loyal customers of the company?



**Answer**: The highly profitable clients are old clients of the company. The highly profitable clients are minimum 16 months with the company and the oldest customer is for the 71 months. Hence, loyal custoers are highly profitable.     
```{r}
claims_df_high_profit_months <- claims_df_high_profit%>% group_by(months_policy_active) %>%count(months_policy_active) 
claims_df_high_profit_months
ggplot(claims_df_high_profit, aes(x = months_policy_active))+
  geom_density(fill="violet", color="#e9ecef", alpha=0.8)+
  ggtitle("High Profitable Customer Duration with company")
  
```
## Question 9



**Question**: Are there types of customers, based on their policy or demographics, that are loss making(negative lifetime value) customers?


**Answer**:
There are 2210 loss making customers. California has highest of such customers. In all the states employed customers bring most loss. Also, personal policy type has most number of loss making customers. Basic coverage is most among the loss making customers. 



```{r}

loss_making_customers 
loss_making_customers%>% group_by(customer_state) %>%count(employment_status)
loss_making_customers%>% group_by(policy) %>%count(policy) 
loss_making_customers%>% group_by(coverage) %>%count(coverage)
```
## Question 10



**Question**: Is there any specific trend between education, residential type, and income of the "problem customers"?



**Answer**: Most of the "problem customers" have completed high schools, associate degree, and bachelor degree. Loss is significantly low for master degree and doctoral. For income between 10k to 40k is the range where we get maximum of the loss, but this continuous till 90k. Most of the "problem customers" are from suburban.

```{r fig.align="center", echo = FALSE,fig.width = 14}
loss_making_customers_education <- loss_making_customers%>% group_by(customer_state) %>%count(highest_education) 
loss_making_customers_education
ggplot(loss_making_customers_education, aes(x=highest_education, y=n, fill = highest_education))+
  geom_col()+
  facet_grid(~customer_state)+
  ggtitle("Problem customers' education in each state ") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r}
loss_making_customers <- loss_making_customers%>%mutate(gross_income = round(income/10000))

ggplot(loss_making_customers, aes(x=factor(gross_income))) +
  geom_histogram(stat="count",fill="red")+
  ggtitle("Problem customers' income distriution ")

```
```{r}
loss_making_customers%>% group_by(residence_type) %>%count(residence_type) 
```
## Question 11

**Question**: Are "problem customers" clients new customers of the company?



**Answer**: The "problem customers" are mostly new clients of the company. The "problem customers" are median 30 months with the company and the oldest customer is for the 69 months and newest 12 months. Hence, "problem customers" are new customers.     
```{r}
loss_making_customers_months <- loss_making_customers%>% group_by(months_policy_active) %>%count(months_policy_active) %>% arrange(months_policy_active)
loss_making_customers_months
loss_making_customers%>%  summarize(median = median(months_policy_active))
ggplot(loss_making_customers, aes(x = months_policy_active))+
  geom_density(fill="violet", color="#e9ecef", alpha=0.8)+
  ggtitle("Problem Customer Duration with company")
  
```


# Summary of Results

Write an executive summary of your overall findings and recommendations to the executives at this company. Think of this section as your closing remarks of a presentation, where you summarize your key findings and make recommendations to improve pricing, company operations, and car insurance policy adjustments.

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm){target="_blank"}, with minimal grammatical errors, and should include the following sections:

1.  An introduction where you explain the business problem and goals of your data analysis

    -   What problem(s) is this company trying to solve? Why are they
        important to their future success?

    -   What was the goal of your analysis? What questions were you
        trying to answer and why do they matter?

<br>

2.  Highlights and key findings from your Exploratory Data Analysis section

    -   What were the interesting findings from your analysis and **why
        are they important for the business**?

    -   This section is meant to **establish the need for your
        recommendations** in the following section

<br>

3.  Your recommendations to the company

    -   Each recommendation must be supported by your data analysis
        results

    -   You must clearly explain **why** you are making each
        recommendation and which results from your data analysis support
        this recommendation

    -   You must also describe the potential business impact of your
        recommendation:

        -   Why is this a good recommendation?

        -   What benefits will the business achieve?
      


## Executive Summary

Please write your executive summary below. If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.

The data belongs to the automobile insurance company. The main problem the company faces is a decrease in profit over the long run. To sustain any insurance company, they must be considered all factors to decide the premium of the policy. 
The goal of my analysis was to identify the factors and customers that are highly profitable to the insurance company and identify any pattern that can be seen in the problem customers. Apart from this, the analysis is done to recommend ways to improve the company’s profitability. I was trying to answer the questions that define the profitability and loss-making features of the customer. Since we are analyzing past data, preventive measures and improvements can be done to improve profitability.

The key findings from the analysis:
1. The gender does not affect the claim rate but married clients have significantly more claim rate.
2. California has significantly higher number of claims. It is followed by Oregon, Arizona, Nevada, and Washington.
3. In all the states employed customers bring most profit. Also, personal policy type has most number of highly profitable customers. Extended coverage is also profitable.
4. Most of the highly profitable customers have completed high schools, associate degree, and bachelor degree. Profitability is significantly low for master degree and doctoral. For income between 10k to 70k is the range where we get most of the profit. Most of the highly profitable customers are from suburban.
5. 1005 problem customers have 3 total claims and 255 customers have 4 total claims. Moreover, there is drastic increase in median loss caused to the company as the total claims increases.
6. Midsize car category is causing more loss for all sales channel. Four-door vehicle class is causing more loss. Also, Agent and branch are the sales channel that has more loss making customers.   
7. Luxury suv and luxury cars are most profitable. All midsize car owners are profitable. Agents and branches are more profitable.
8. The highly profitable clients are old clients of the company.


The goal is to increase the profit and decrease loss-making. Here are a few recommendations for this:
1.Increase the premium amount for married clients.  The number of claims made by them is highest compared to single and divorced clients.
2.California state has the highest number of claims, followed by Oregon.  It would be better if the company could increase the premium amount in these states.
3.The company has more “Personal” policy types with extended coverage with the most profit. If a possible company could expand to accommodate new corporate clients. The most loss is seen with basic coverage. The company can review the rates for coverage so that clients find “extended” coverage more beneficial.
4.For all the states, master and doctoral clients have low drastic claim rates. The company can give more discounts to the clients with such high education, to attract new and preserve the old clients.
5.Suburban has the most profitable and loss-making customers. We can increase the premium amount to increase the profit, apart from this we can have discounts for rural and urban residents to attract new customers.
6.The company should monitor the claims made by the customers. If they have done more than 2 claims, the company should increase their premium amount.
7.Agent and branches are attracting both the most loss-making customers and profitable customers. The company should create some training programs to improve the quality of agents and branch employees so that they can access new clients and give them a premium amount that is beneficial for the company.
8.Loyal customers are highly profitable. It would be better if the company look into keeping the old clients first. To keep the clients for longer, the company could provide discounts to clients as they remain the client.















