
```{r warning = FALSE, message = FALSE}
# Suppress dplyr summarise grouping warning messages
options(dplyr.summarise.inform = FALSE)

## Add R libraries here
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(skimr)

# Load the dataset
telecom_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/telecom_df.rds'))

```
# Data Analysis
Explore the dataset:
unique values in data:
```{r}
skim(telecom_df)
```
checking summary for numeric values:
```{r}
summary(telecom_df$avg_call_mins)
```
```{r}
summary(telecom_df$months_with_company)
```
```{r}
summary(telecom_df$monthly_charges)
```
```{r}
sum(is.na(telecom_df))
summary(telecom_df$avg_intl_mins)
```
# Question 1


**Question**: Are we losing old clients?


**Answer**: No, from the boxplot below we can say, new customers are canceling the service. The customers who are canceling the service are staying for lesser tenure with the company. The cancelling customers have tenure of median 10 months with the company, whereas the customer continuing the service have tenure of median 40 months. Hence, we are not losing old clients but rather the new ones.


```{r}
telecom_df %>%
ggplot(data = telecom_df, mapping = aes(x = canceled_service, y= months_with_company,fill = canceled_service))+ 
  geom_boxplot()+
  labs(title = "Months with company with current service status",
       x = "Canceled serivce Status",
       y = "Months with company")


```



# Question 2


**Question**: Does family status (spouse/ partner and dependents) affect the cancellation rate?


**Answer**: There is less cancellation in people with spouse and dependents than the ones without it. From the below summarized data and plot,
customers with spouse/partner and customers with dependents cancel the service less often than the customers without them. 


```{r}
family_status <- telecom_df %>% 
  group_by(canceled_service, spouse_partner, dependents) %>%
  summarize(Count = n()) %>%
  arrange(canceled_service)
family_status

ggplot(data = family_status, mapping = aes(x=spouse_partner, y = Count, fill = canceled_service)) +
  geom_col(position = "fill")+
  labs(title = "Customer retention based spouse/partner",
       x = "Spouse / Partner",
       y = "Proportion")

```
```{r}
ggplot(data = family_status, mapping = aes(x=dependents, y = Count, fill = canceled_service)) +
  geom_col(position = "fill")+
  labs(title = "Customer retention based dependents",
       x = "Dependents",
       y = "Proportion")

```

# Question 3


**Question**: Does longer contract mean retention of the customers?


**Answer**: Yes, longer contracts do mean retention of the customer. Month-to-month contract holders have the more proportion of cancellation of service compared to the other two. Also, customers with one year contract leave the company more than the customers with two year contract. 


```{r}
contract_status <- telecom_df %>% 
  group_by(canceled_service, contract) %>%
  summarize(Count = n())

ggplot(data = contract_status, mapping = aes(x = canceled_service, y= Count, fill=contract))+
   geom_bar(position="fill", stat = "identity")+
  labs(title = "Customer retention based on contract type",
       x = "Canceled Service",
       y = "Proportion")


```

# Question 4


**Question**: Does monthly charges affect the cancellation of service?


**Answer**: As can be seen from the graph below, there is significant increase in the cancellation as the monthly charges are increasing. 


```{r}
                    
ggplot(data = telecom_df, aes(x = monthly_charges, fill = canceled_service)) +
  geom_histogram(aes(y = ..density..), color = "grey", bins = 10) +
  facet_wrap(~ canceled_service, nrow = 2) +
  labs(title = "Effect of monthly charges on cancellation", x = "Monthly Charges", y = "Proportion")



```



# Question 5


**Question**: Does cellular service type has any affect on the cancellation of service? 


**Answer**: Customers with multiple lines tend to cancel the service less.


```{r}
services_df <- telecom_df %>%
                group_by(canceled_service,cellular_service) %>%
                summarize(Count = n(), median_expense = median(avg_call_mins))

services_df

ggplot(data = services_df, mapping = aes(x=canceled_service, y=Count, fill = canceled_service))+
  geom_col()+
  facet_grid(~cellular_service)+
  labs(title ="Cellular service with canceled service")
```
# Question 6


**Question**: Does more availed services result in retention of the customers?


**Answer**: For this analysis, we considered the total services used by customers. Customers with more using multiple services are not leaving the company service as much as the customers with lesser services.




```{r}
old_value <- c("yes","no")
new_value <- c(1,0)               
new_telecom_df <- telecom_df %>% select(canceled_service,online_security,online_backup,device_protection,tech_support,streaming_tv,streaming_movies)

new_telecom_df$online_security <- as.character(new_telecom_df$online_security)
new_telecom_df$online_security[new_telecom_df$online_security == "yes"] <- 1
new_telecom_df$online_security[new_telecom_df$online_security == "no"] <- 0

new_telecom_df$online_backup <- as.character(new_telecom_df$online_backup)
new_telecom_df$online_backup[new_telecom_df$online_backup == "yes"] <- 1
new_telecom_df$online_backup[new_telecom_df$online_backup == "no"] <- 0

new_telecom_df$device_protection <- as.character(new_telecom_df$device_protection)
new_telecom_df$device_protection[new_telecom_df$device_protection == "yes"] <- 1
new_telecom_df$device_protection[new_telecom_df$device_protection == "no"] <- 0

new_telecom_df$tech_support <- as.character(new_telecom_df$tech_support)
new_telecom_df$tech_support[new_telecom_df$tech_support == "yes"] <- 1
new_telecom_df$tech_support[new_telecom_df$tech_support == "no"] <- 0

new_telecom_df$streaming_tv <- as.character(new_telecom_df$streaming_tv)
new_telecom_df$streaming_tv[new_telecom_df$streaming_tv == "yes"] <- 1
new_telecom_df$streaming_tv[new_telecom_df$streaming_tv == "no"] <- 0

new_telecom_df$streaming_movies <- as.character(new_telecom_df$streaming_movies)
new_telecom_df$streaming_movies[new_telecom_df$streaming_movies == "yes"] <- 1
new_telecom_df$streaming_movies[new_telecom_df$streaming_movies == "no"] <- 0
new_telecom_df
```
```{r}
new_telecom_df <- new_telecom_df %>% mutate(total_services = as.numeric(online_security) + as.numeric(online_backup) + as.numeric(device_protection) + as.numeric(tech_support) + as.numeric(streaming_tv) + as.numeric(streaming_movies))

new_telecom_df
```
```{r}
ggplot(data = new_telecom_df, aes(x = total_services, fill = canceled_service)) +geom_histogram(stat = "count", color = "grey", bins = 7) + facet_wrap(~ canceled_service, nrow = 2) + labs(title = "Effect of number of services on cancellation of service", x = "Total services used", y = "Count")
```
# Question 7


**Question**: Are we losing senior citizen customers?


**Answer**: Yes, we are losing senior citizen customers. From the data below it is clear that number of senior citizen who have canceled the service is almost same as the ones who did not.
```{r}

senior_citizen_df <- telecom_df %>% 
                    filter(senior_citizen == "yes") %>%
                    group_by(canceled_service) %>%
                    summarize(Count = n())
senior_citizen_df 
```




# Machine Learning

```{r}
set.seed(101)

#data splitting and feature engineering steps

telecom_split <- initial_split( telecom_df,prop = 0.70,strata = canceled_service)
telecom_training <- telecom_split %>% training()
telecom_test <- telecom_split %>% testing()
```

```{r}
set.seed(101)

#5-fold cross validation
telecom_folds <- vfold_cv(telecom_training, v = 5)
```

```{r}
telecom_recipe <- recipe(canceled_service ~., data = telecom_training) %>% 
                 step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
                 step_normalize(all_numeric(), -all_outcomes()) %>% 
                 step_dummy(all_nominal(), -all_outcomes())

telecom_recipe %>% prep(training=telecom_training) %>% bake(new_data = NULL)
telecom_test

```



# Model 1 - Logistic Regression Model
```{r}
logistic_model <- logistic_reg() %>% 
                  set_engine('glm') %>% 
                  set_mode('classification')

# Create a workflow
logistic_workflow <- workflow() %>% 
               add_model(logistic_model) %>% 
               add_recipe(telecom_recipe)

# Fit model
logistic_fit <- logistic_workflow %>% 
                last_fit(split = telecom_split)

# Collect predictions
logistic_results <-  logistic_fit %>% 
                     collect_predictions()
logistic_fit %>% collect_metrics()
```

Evaluationg the model performance:
```{r}
## ROC Curve
roc_curve(logistic_results, 
          truth = canceled_service, 
          estimate = .pred_yes) %>% 
  autoplot()
```
```{r}
# Confusion Matrix
conf_mat(logistic_results, 
         truth = canceled_service, 
         estimate = .pred_class)
```

```{r}
telecom_metrics <- metric_set(accuracy, sens, spec, f_meas, roc_auc)
telecom_metrics(logistic_results, truth = canceled_service,
                estimate = .pred_class, .pred_yes)
```

# Model 2 - KNN

```{r}
library(kknn)
knn_model <- nearest_neighbor(neighbors = tune()) %>% 
             set_engine('kknn') %>% 
             set_mode('classification')
```

```{r}
knn_workflow <- workflow() %>% 
          add_model(knn_model) %>% 
          add_recipe(telecom_recipe)
```

```{r}
k_grid <- tibble(neighbors = c(10, 15, 25, 45, 60, 80, 100, 120, 140, 180))

set.seed(101)

knn_tuning <- knn_workflow %>% 
              tune_grid(resamples = telecom_folds,
                        grid = k_grid)
```



```{r}
knn_tuning %>% show_best('roc_auc')
```


```{r}
## Select best model based on roc_auc
best_k <- knn_tuning %>% 
          select_best(metric = 'roc_auc')

best_k
```
```{r}
## Finalize workflow by adding the best performing model

final_knn_wf <- knn_workflow %>% 
                finalize_workflow(best_k)
```

```{r}
last_knn_fit <- final_knn_wf %>% 
           last_fit(split = telecom_split)

last_knn_fit %>% collect_metrics()
```

```{r}
knn_predictions <-  last_knn_fit %>% 
                collect_predictions()
knn_predictions
```
```{r}
## ROC Curve
knn_predictions %>%
roc_curve(truth = canceled_service, 
          estimate = .pred_yes) %>% 
  autoplot()
```
```{r}
# Confusion Matrix
conf_mat(knn_predictions, 
         truth = canceled_service, 
         estimate = .pred_class)
```

```{r}
telecom_metrics2 <- metric_set(accuracy, sens, spec, f_meas, roc_auc)
telecom_metrics2(knn_predictions, truth = canceled_service,
                estimate = .pred_class, .pred_yes)
telecom_metrics2


```





# Model 3 - Random forest

```{r}
library(vip)
library(rpart.plot)

randomf_model <- rand_forest(mtry = tune(),trees = tune(),min_n = tune()) %>% 
                 set_engine('ranger', importance = "impurity") %>% 
                 set_mode('classification')

```

```{r}
random_workflow <- workflow() %>% 
          add_model(randomf_model) %>% 
          add_recipe(telecom_recipe)
```

```{r}
set.seed(101)

randomf_grid <- grid_random(mtry() %>% range_set(c(2, 5)),trees(),min_n(),size = 10)

randomf_grid

set.seed(101)

randomf_tuning <- random_workflow %>% tune_grid(resamples = telecom_folds,grid = randomf_grid)
```

```{r}
randomf_tuning %>% show_best('roc_auc')
```
```{r}
## Select best model based on roc_auc
best_random <- randomf_tuning %>% 
          select_best(metric = 'roc_auc')

best_random
```
```{r}
## Finalize workflow by adding the best performing model

final_random_wf <- random_workflow %>% 
                finalize_workflow(best_random)
```

```{r}
random_wf_fit <- final_random_wf %>% 
           fit(data = telecom_training)
```

```{r}
random_fit <- random_wf_fit %>%
              extract_fit_parsnip()
```

```{r}
vip(random_fit)
```
```{r}
random_last_fit <- final_random_wf %>% 
               last_fit(telecom_split)
```


```{r}
random_last_fit %>% collect_metrics()
```
```{r}
random_predictions <-  random_last_fit %>% 
                collect_predictions()
random_predictions
```
```{r}
## ROC Curve
random_predictions %>%
roc_curve(truth = canceled_service, 
          estimate = .pred_yes) %>% 
  autoplot()
```
```{r}
# Confusion Matrix
conf_mat(random_predictions, 
         truth = canceled_service, 
         estimate = .pred_class)
```
```{r}
telecom_metrics3 <- metric_set(accuracy, sens, spec, f_meas, roc_auc)
telecom_metrics3(random_predictions, truth = canceled_service,
                estimate =.pred_class, .pred_yes)
```


**Summary**
Introduction:

Our data is from a telecom company. The profit of the company depends on the acquisition of new customers as well as the retention of old customers. The company wants to find out the factors that are responsible for customers canceling the service. There can be multiple reasons involved that would be responsible for clients leaving the company, e.g.., bad service, better prices from competitors, etc. This analysis provides machine learning algorithms to predict the possibility of the customer leaving and identify different factors that are responsible for a customer leaving.
The goal of the analysis is to help the company to help in identifying the pattern of customers that are leaving the company and predict whether a customer might leave the company. The analysis will provide insight so that it can take measures to reduce customers leaving the company service.

Key findings from the analysis:
1.	New customers are canceling the service more. The customers who are canceling the service are staying for lesser tenure with the company. The canceling customers have a tenure of a median of 10 months with the company, whereas the customer continuing the service have a tenure of a median of 40 months. Hence, we are not losing old clients but rather new ones.
2.	There are fewer cancellations in people with spouses and/or dependents than the ones without them. 
3.	Longer contracts mean retention of the customer. Month-to-month contract holders have a more proportion of cancellations of service compared to the other two. Also, customers with one-year contracts leave the company more than customers with two-year contracts.
4.	Customers with more monthly charges are leaving the company more. 
5.	Customers with a single line tend to cancel the service more.
6.	The more services availed by the customer, the more the chances of retention. 
7.	Senior citizens are leaving the company services at an alarming rate.

Model analysis:
I have used three different classification machine learning algorithms to classify possible customers who will leave the company’s service. I have implemented logistic regression, KNN, and random forest models. 
For the business, the model must have a lower false negative rate.  It is important for the business to retain its customers. So, we should be able to identify the customers who might cancel the service as soon as and as accurately as possible. We achieve this by taking into consideration the model’s sensitivity. 
From my observation, KNN is the best model. It gives more sensitivity value compared to the other two models. It is sensitivity value is approx. 0.76 and it only gave 31 (which is the least among all the models) false negatives.



Recommendations:

Add your summary here. Please do not place your text within R code chunks.
After carefully analyzing the data, I have the following recommendations for the company:
1.	The company is not losing old clients. The company should give some attractive benefits/discounts to the new customers. It is also seen that the company is having difficulties in retaining customers with lesser months.
2.	The company should have new economic plans for customers with no spouse(partner) and/or dependent. This will attract new single customers and retain them. Customers with spouses and/or dependents are less likely to terminate the service. 
3.	The company should introduce new lucrative longer plans. Currently, the company has only 3 plans, (monthly, yearly, and two-yearly). The company could introduce new longer plans with 3 months, 6 months, 9 months, 15 months, and so on. This will help the company to retain its customers.
4.	 The company could give some discounts/offers to its expensive customers. The customers with more monthly charges would stay with the company if they get the service for cheaper or more value is added to their plans.
5.	The company should introduce new plans or offer discounts or give new benefits to the customers with a single line. This will retain the customers with single-line service. 
6.	The company can offer service bundles to its customers at a discounted rate. The more services customer avail lesser the chances they will leave the company’s service.
7.	Senior citizens should get some benefits to retain them as a customer. We are losing senior citizen customers at an alarming rate. We should devise some strategy to stop this.


